---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Sophie Boyd"
date: "2-6-26"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r dennys-dimensions}
view(dennys)
nrow(dennys)
ncol(dennys)
```

The Denny's dataset contains 1643 observations of 6 variables. Each row represents one Denny's establishment. The variables are address, city, state, zip code, longitude, and latitude.

### Exercise 2

```{r laquinta-dimensions}
view(laquinta)
nrow(laquinta)
ncol(laquinta)
```

The La Quinta dataset contains 909 observations of 6 variables. Each row represents one La Quinta establishment. The variables are the same as the Denny's dataset: address, city, state, zip code, longitude, and latitude.

### Exercise 3

...

### Exercise 4

...

### Exercise 5

...

### Exercise 6

...

 
Add exercise headings as needed.

